---
title: "SCINET demo (basic functionality)"
output:
  html_document:
    df_print: paged
---

## Housekeeping

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
require(SCINET)
require(ACTIONet)

```


## Preprocessing using ACTIONet framework
```{r message=FALSE, warning=FALSE}
sce = readRDS('reduced_sce_PBMC_annotated.RDS')

ACTIONet.out = readRDS('ACTIONet_out_PBMC.RDS')

arch.annot = annotate.archetype(ACTIONet.out, sce$Labels)

A = ACTIONet.out$signature.profile

```

## Read marker genes
```{r}
marker.genes = readRDS('human_pbmc_marker.RDS')
positive.genes = sapply(marker.genes, function(gs) {
  genes = intersect(rownames(sce), stringr::str_replace(gs[grep("+", gs, fixed = TRUE)], stringr::fixed("+"), ""))
})
```


## Visualize unnormalized gene activity scores
```{r}
  genes = sort(unique(unlist(positive.genes)))
  genes.idx = match(genes, rownames(sce))

  sub.profile = A[genes.idx, ACTIONet.out$core.out$core.archs]
  # sub.profile[sub.profile > 5] = 5
  
  rownames(sub.profile) = genes
  colnames(sub.profile) = arch.annot$archetypeLabels[ACTIONet.out$core.out$core.archs]
  
  
  Annot = levels(sce$Labels)
	colCelltype.Pal = ggpubr::get_palette("d3", length(Annot))
	names(colCelltype.Pal) = Annot
	
	require(ComplexHeatmap)
  
  ha_cols = HeatmapAnnotation(df = list(Celltype = colnames(sub.profile)), col = list(Celltype = colCelltype.Pal), annotation_legend_param = list(Celltype=list(title_gp = gpar(fontsize = 8), labels_gp = gpar(fontsize = 5))), which = "column")  
  
  gradPal = grDevices::colorRampPalette(rev(RColorBrewer::brewer.pal(n = 9, name = "RdYlBu")))(100)

  Heatmap(sub.profile, col = gradPal, row_names_gp = gpar(fontsize = 6), column_names_gp = gpar(fontsize = 0), top_annotation = ha_cols, name = "Expression of marker genes", row_title = "Genes", column_title = "Cell states")
  
```


## Estimate gene activity score within B-cell sub-population
```{r}
# It can be any imputed profile

require(SCINET)
activity.scores = SCINET::compute_gene_activities_full(A = A, thread_no = 8)

rownames(activity.scores) = rownames(sce)
colnames(activity.scores) = arch.annot$archetypeLabels

```


## Visualize normalized gene activity scores
```{r}
  genes = sort(unique(unlist(positive.genes)))
  genes.idx = match(genes, rownames(sce))

  sub.profile = activity.scores[genes.idx, ACTIONet.out$core.out$core.archs]

  Annot = levels(sce$Labels)
	colCelltype.Pal = ggpubr::get_palette("d3", length(Annot))
	names(colCelltype.Pal) = Annot
	
	require(ComplexHeatmap)
  
  ha_cols = HeatmapAnnotation(df = list(Celltype = colnames(sub.profile)), col = list(Celltype = colCelltype.Pal), annotation_legend_param = list(Celltype=list(title_gp = gpar(fontsize = 8), labels_gp = gpar(fontsize = 5))), which = "column")  
  
  gradPal = grDevices::colorRampPalette(rev(RColorBrewer::brewer.pal(n = 9, name = "RdYlBu")))(100)

  Heatmap(sub.profile, col = gradPal, row_names_gp = gpar(fontsize = 6), column_names_gp = gpar(fontsize = 0), top_annotation = ha_cols, name = "Expression of marker genes", row_title = "Genes", column_title = "Cell states")
  
```



## Load GAAD network as the reference interactome 
Downloaded from [GAAD](http://gaad.medgenius.info/Downloads/co-occurrence_of_gene_pairs_in_ADs.txt.gz)

```{r}
GAAD = readRDS('GAAD_net.RDS')
paired.datasets = pair.datasets(GAAD, activity.scores)

EL = get.edgelist(GAAD, names = FALSE)

G.adj = as(get.adjacency(paired.datasets$net), 'dgTMatrix')

edge.idx = (G.adj@i)*nrow(G.adj) + (G.adj@j+1)

```



```{r}

nets = construct_cell_networks_noPrior(gene_activities = paired.datasets$activity.scores[, 1:10], thread_no = 8)


```


## Compute topological-specificity of genes in each network
```{r}
require(igraph)

# core.topo.spec = sapply(cellstate.nets[ACTIONet.out$core.out$core.archs], function(Adj) {
core.topo.spec = sapply(nets, function(Adj) {
  G = graph_from_adjacency_matrix(Adj, mode = "undirected", weighted = TRUE)
  V(G)$name = paired.datasets$genes
  
  topo.spec = topo.spec(G, sample_no = 1000)

  return(topo.spec)
})

  colnames(core.topo.spec) = arch.annot$archetypeLabels[1:10]#ACTIONet.out$core.out$core.archs]
  rownames(core.topo.spec) = paired.datasets$genes
  
```


## Constrained networks
```{r}
cellstate.nets = construct_cell_networks(net = G.adj, gene_activities = paired.datasets$activity.scores, thread_no = 8)

```

